---
title: "DataGeneration"
author: "Eric Massa"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

R Markdown document "DataGeneration.Rmd" for BIOL 812 week 3 assignment. Each step is labelled with comments in the code below. Additional commentary is included as needed.

Steps were as follows:


1.    Create a new public repository on your GitHub account called USERNAME.github.io (but replace USERNAME with your own GitHub username). This will create a public website that anyone can access with a web browser, by entering USERNAME.github.io (it may take a few minutes to sync).
2.    Now open RStudio and create a new project linked to your USERNAME.github.io directory. Be sure to use the repository website, not the public website link: i.e. http://github.com/USERNAME/USERNAME.github.io NOT http://USERNAME.github.io
3.    Create a new Rmd script called DataGeneration.R Inside of DataGeneration.R
4.    Create an object called Species that is a character vector containing the names of 10 species (you can use any species names you want)
5.    Create a vector called Limb.Width that is a numeric vector containing 10 random numbers sampled from a normal distribution with a mean of 10 and standard deviation of 10.
6.    Change any negative numbers in Limb.Width to positive values. Make sure this will work regardless of which random numbers are generated in 2a.
7.    Create a vector called unitsW that is a boolean vector containing 10 strings that are either mm or cm generated randomly with a 0.5 probability of each
8.    Repeat step 5 and 6 for a new object called Limb.Length, but with a mean and standard deviation of 100
9.    Repeat step 7 for a new object called unitsL
10.    Combine all vectors into a data.frame object called MyData
11.    Save MyData as a csv file in the Data folder called measurements.csv


```{r }
#Part 1, steps 1-3 involve creation of repositories/files, not code
#Part 1, step 4
Species <- c("Ursus americanus","Ursus arctos","Ursus ursinus","Ursus maritimus","Ursus malayanus","Ursus thibetanus","Ursus deningeri","Ursus etruscus","Ursus minimus","Ursus rossicus")
```

## Start of data creation
Create a vector of 10 species names.

```{r }
#Part 1, step 5
Limb.Width <- rnorm(n = 10, mean = 10, sd = 10)
#Part 1, step 6
for (i in 1:length(Limb.Width)) {
  if ((Limb.Width[i] < 0) == TRUE) {
    Limb.Width[i] <- -1*Limb.Width[i]
  }
}
```
Generate 10 limb width values with a mean value of 10 and a standard deviation of 10.
If a value is negative, multiply by -1 to make it positive.
```{r }
#Part 1, step 7
unitsW <- sample(x = c("cm","mm"),size = 10, replace = TRUE, prob = c(0.5,0.5))
```

Randomly assign each limb width value a unit of either millimeters (mm) or centimeters (cm).

```{r }
#Part 1, step 8
Limb.Length <- c(rnorm(n = 10, mean = 100, sd = 100))
for (i in 1:length(Limb.Length)) {
  if ((Limb.Length[i] < 0) == TRUE) {
    Limb.Length[i] <- -1*Limb.Length[i]
  }
}
#Part 1, step 9
unitsL <- sample(x = c("cm","mm"),size = 10, replace = TRUE, prob = c(0.5,0.5))
```

Generate 10 limb length values with mean value of 100 and standard deviation of 100 and make sure that value is positive.
Multiply by -1 to make it positive if needed.
Randomly assign unit of either millimeters (mm) or centimeters (cm).

```{r }
#Part 1, step 10
MyData <- data.frame(Species, Limb.Width, unitsW, Limb.Length, unitsL)
#Part 1, step 11
write.csv(MyData, file = "measurements.csv",row.names = FALSE)
```
Create dataframe of variables generated and save to "measurements.csv" file.