---
title: "DataAnalysis"
author: "Eric Massa"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

R Markdown document "DataGeneration.Rmd" for BIOL 812 week 3 assignment. Each step is labelled with comments in the code below. Additional commentary is found below as needed.

Steps were as follows:

1.    Read measurements.csv into an object in R called MData
2.    Include functions from the Data Science Tutorial to do each of the following:
+    Change the column names to shorter, 3-letter abbreviations
+    Add a new data column called Volume that calculates limb volume (Hint: pay careful attention to your units – assume unitsW shows the units for Limb.Width and unitsL shows the units for Limb.Length). Treat these as limb dimensions (e.g. arms, legs, fins).
+    Group the species into 2 or more imaginary groups (you decide on the category names) and add a column denoting these groups
+    Summarize the mean and SD of Length, Width and Volume for each of your artificial groups, and present in table form.
+    Plot Width and Length, as in Part III from last week’s assignment BUT embed the graph in your R markdown file rather than creating a separate pdf.
+    Plot Width and Length, but only for species that have above-average volume
+    Add a table showing only Species and Volume for only the species with below-average volume
3.    Add notes to your ‘official’ report.
4.    Knit your Rmd file to a html format.
5.    Save and upload your files to the corresponding GitHub repository (hint: review ‘commit’, ‘push’ and ‘pull’)
6.    Submit a link to your GitHub repository to OnQ

```{r}
#Necessary libraries
library(dplyr)
library(ggplot2)
```

##Read data

```{r }
#Part 2, step 1
MData <- read.csv("measurements.csv")
```

##Rename dataframe columns

```{r }
#Part 2, step 2, bullet 1
MData2 <- rename(.data = MData, SPP = Species, LWD = Limb.Width, LLG = Limb.Length, UNW = unitsW, UNL = unitsL)
```

##Calculate limb volume
First convert all limb dimensions into cm.
Second, calculate limb volume in cm^3
```{r }
#Part 2, step 2, bullet 2
cmL <- numeric()
cmW <- numeric()
for (i in 1:nrow(MData2)) {
  if (MData2$UNL[i] == "mm") {
    cmL[i] <- MData2$LLG[i]/10
  } else {
    cmL[i] <- MData2$LLG[i]
  }
}
for (i in 1:nrow(MData2)) {
  if (MData2$UNW[i] == "mm") {
    cmW[i] <- MData2$LWD[i]/10
  } else {
    cmW[i] <- MData2$LWD[i]
  }
}
vol <- numeric()
for (i in 1:10) {
  vol[i] <- pi * cmW[i]^2 * cmL[i]
}
```

append volume data to imported dataset.

```{r }
MData3 <- mutate(MData2, Volume = vol)
```

## Data grouping and summary
Arbitrary grouping of data into groups "A" and "B".
Summarize limb data (length, width, and volume), presenting mean and standard deviation of each (in cm) by group.

```{r }
#Part 2, step 2, bullet 3
MData4 <- mutate(MData3, Group = c(rep("A",5), rep("B",5)))
#Part 2, step 2, bullet 4
summarize(.data = group_by(MData4, Group), MeanL = mean(cmL), SDL = sd(cmL), MeanW = mean(cmW), SDW = sd(cmW), MeanV = mean(Volume), SDV = sd(Volume))
```

##Plots
Plots demonstrating relationships in the data.
The first plot demonstrates the relationship between limb length and limb width of all organisms in the data.
```{r }
#Part 2, step 2, bullet 5
qplot(x = cmL, y = cmW, xlab = "Length of limb (cm)", ylab = "Width of limb (cm)")
```

The second plot demonstrates the same, but only for those species with limb volumes above average.

```{r }
#Part 2, step 2, bullet 6
qplot(x = cmL[MData4$Volume > mean(MData4$Volume)], y = cmW[MData4$Volume > mean(MData4$Volume)], xlab = "Length of limb (cm)", ylab = "Width of limb (cm)")
```

Show species name and limb volume for species with below-average limb volume.

```{r }
#Part 2, step 2, bullet 7
filter(.data = MData4[,c(1,6)], Volume < mean(MData4$Volume))


#Part 2, step 3
#...

#Part 2, step 4
#Knit to file, no code.
#Part 2, step 5
#Saving and uploading files to GitHub repository. No code.
#Part 2, step 6
#Submit to OnQ. No code.
```
