---
title: "Regex"
author: "Eric Massa"
date: "1/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Code below does the following
1. Import measurements.csv into an object called MData
2. Use a dplyr command AND regular expressions to add a new column called Sp that modifies the Species names with only the first letter and period.
3. Use a dplyr command to create a new dataset BUT with Sp column instead of the Species column, moved to the beginning in place of the original Species column

Comments within code state what the line of code immediately following does.


```{r}
#Import library 'dplyr'
library(dplyr)

# Import measurements.csv file into object MData
MData <- read.csv("measurements.csv")

#Create column Sp, consisting of genus initial and species name, using dplyr and regex
MData <- mutate(MData, Sp = sub("^(\\w)\\w{2,}", "\\1\\.", x = MData$Species))

#Print contents of MData
MData

#Create a new dataset (MDataAbbrev) with Sp column at beginning of dataset in place of original Species column.
MDataAbbrev <- select(MData, -Limb.Width:-unitsL, everything()) %>%
  select(-Species)

#Print contents of MDataAbbrev
MDataAbbrev

```
